---
title: Twitter Vaccination Sentiment Analysis
author: "Alexandra Stephens"
output: 
    html_document:
        theme: cosmo 
        toc: true
        toc_float: true
        highlight: tango
        number_sections: false
---

```{r setup, include=FALSE, warnings=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Motivation
It is important to identify communities who are at higher risk of infectious diseases in order to implement preventative measures. In this document, we will classify Twitter users and their tweets as "pro" or "anti" vaccination. 

This code consists of data exploration and visualization using Natural Language Processing techniques, a social network analysis of users with the `igraph` package, and a sentiment analysis of the results.

The libraries used in this study are `stringr`, `tidytext`, `dplyr`, `ggplot2`, `wordcloud`, `SnowballC`, `igraph`, `threejs`, and `visNetwork`. In order to run this code please ensure you have these packages installed. If you would like to download your own Twitter data, you will need to sign up for mma Twitter developer account, create an application, and enter your own keys & credintials in the "What is the data?" section below.

The learning objectives include data cleaning and manipulation with natural language processing & regular expressions, social network analysis, semi-supervised learning concepts, and sentiment.

# What is the data?
To collect data from Twitter, we suggest using the `rtweet` package. Again, if you would like to download your own Twitter data, you will need to sign up for a Twitter developer account, create an application, and enter your own keys below.

```{r eval=FALSE}
library(rtweet)

twitter_token <- create_token(
  app = "twitter_app_name",
  consumer_key = "XXXXXXXXXXXXXXX",
  consumer_secret = "XXXXXXXXXXXXXXXXXXXXX",
  access_token = "XXXXXXXXXXXXXXXXXXXXXXXXXXX",
  access_secret = "XXXXXXXXXXXXXXXXX")
```
Once the twitter_token is create it is stored in the workspace and you are free to use all the functions in `rtweet`!

The `search_tweets` function can download 18,000 every 15 minutes (with the free developer account). However adding `retryonratelimit = TRUE` automatically waits for the rate limit to be reset (after 15 min) and then searches for the remaining tweets. This is repeated until all requested tweets are downloaded.

Below, we are searching for 18,000 recent tweets that contain the hashtags "antivax" or "vaccineswork", as well as just the word "vaccines".

```{r eval=FALSE}
st <- search_tweets('#antivax OR vaccines OR #vaccineswork', 
                    n = 18000,
                    type = "recent",
                    include_rts = TRUE, 
                    lang = "en")
```

# Data import
First we import the libraries needed, then upload the data with `read.csv`.
Make sure you have the libraries below installed before running the code.
```{r, message = FALSE}
library(kableExtra)
library(stringr)

library(tidytext)
library(dplyr)
library(ggplot2)
library(wordcloud)
library(SnowballC)

library(igraph)
library(threejs)
library(visNetwork)

tweets_df <- read.csv("data/date01_14_n18k_key2_uuidgens_final.csv",stringsAsFactors = FALSE)
```


# Data wrangling
## Cleaning with regular expressions
Using regular expressions to parse and clean text is an important skill when dealing with text data. Let's take a look at the data to see what we need to clean.

```{r}
head(tweets_df$text)
```

We want to remove text that does not aid our analysis and may in fact distract from our analysis. For example, we observe lots of links that begin with "https..." that aren't giving us any insightful information.

The string `pat` below contains several regular expression or "regex" patterns that match with a few different unwanted characters or phrases. The "|" character seperates each expression, and when `str_replace_all` is called it checks for matches with every pattern in `pat`. `str_replace_all` also looks for repeats of any pattern, so if there are multiple links it will replace all of them with an empty space, `""`.

|Regex Phrase|Text match|
|---|---|
|`[\r\n]`|Newline character|
|`&amp.`|"&amp" phrase|The "." matches any character|
|`@.*?\\S+`|Tagged user handle followed by whitespace|
|`@.*?$`|Tagged user handle at the end of the tweet|
|`https?:.*\\S+`|Embedded link followed by whitespace|
|`https?:.*$`|Embedded link at the end of the tweet|


```{r}
pat <- "[\r\n]|&amp.|@.*?\\S+|@.*?$|https?:.*\\S+|https?:.*$"

tweets_df$text <- 
  tweets_df$text %>%
    str_replace_all(pattern = pat, "")

tweets_df$retweet_text <- 
  tweets_df$retweet_text %>%
    str_replace_all(pattern = pat, "")

head(tweets_df$text)
```

## The tidy text and NLP

The tidy text format is using **one-token-per-row** in our dataframe. In short, tokens in natural language processing are words. To be technical, you could say a token is a list of characters between two spaces. Tokenization is the process of breaking up a string, or paragraph or document into word bits.

So we want to tokenize our tweets, however we don't want to lose any information. Specifically, we want to keep track of what Tweet each word came from since our analysis involves differentiating tweets. Other analyses including topic summarization may not require keeping track of this--you can just a create a jumble (er, "corpus") of words and find commonalities.

Thus we add the `int` column to the original dataframe `text_df` as an index for the tweets. We then tokenize the text data with tidy text `unnest_tokens` and create a new dataframe `text_df`.

```{r}
text_df <- tweets_df %>% select("text")
text_df$int <- c(1:length(text_df$text))
text_df <- text_df %>%
    unnest_tokens(word, text)
head(text_df)
```

Next we want to remove stop words from the text. Stop words in NLP are common words that don't add value to a sentence, paragraph, document etc. For example, "still think the flu shot is" doesn't give us any extra information than "still think flu shot," so we can remove these words to make our dataset more concise.

We load the `stop_words` dataset, which has two columns: `word` and `lexicon`. There are three different lexicons that reference the *source* of the stop word: ["onix"](http://www.lextek.com/manuals/onix/stopwords1.html), ["SMART"](http://www.jmlr.org/papers/volume5/lewis04a/lewis04a.pdf), and ["snowball."](http://snowball.tartarus.org/algorithms/english/stop.txt) We count the number of stop words in each category so that we can modify this if needed. (For example, if by removing all possible stop words we aren't left with enough information, we could instead just remove the 174 `snowball` stopwords).

```{r}
data(stop_words)

stop_words %>% 
    group_by(lexicon) %>% 
    tally()

text_df <- text_df %>%
  anti_join(stop_words)

head(text_df,10)
```
##Word stemming
Another technique often used alongside removing stopwords is token *stemming*. Stemming changes each word to a root word so that different conjugations of the same word are not treated as different words completely. A human can tell that "vaccinate" and "vaccination" have similar meaning.On the computer, if we were to count every repetition of "vaccinate" we would not get an accurate total since "vaccination" would be ignored.

We will use the library `SnowballC` which implements the [Porter stemming algorithm](http://snowball.tartarus.org/algorithms/porter/stemmer.html).

```{r}
text_df <- text_df %>%
  mutate(word = wordStem(word))
head(text_df,10)
```

# Exploratory data analysis

```{r echo = FALSE}
anti_lst <- as.numeric(
  grepl("antivax", tweets_df$text))
pro_lst <-as.numeric(
  grepl("vaccineswork", tweets_df$text))

tweets_df$Hashtags <- "Neither"
tweets_df$Hashtags[anti_lst == 1] <- "#antivax"
tweets_df$Hashtags[pro_lst == 1] <- "#vaccineswork"
```


```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(ggmap)

states <- map_data("state")

ggplot(data = states) + 
  geom_point(data = tweets_df, aes(x = lng, y = lat,  colour = Hashtags), size = 1) +
  geom_polygon(aes(x = long, y = lat, group=group), fill = NA, color = "dark grey") +
  theme_void()

sum(is.na(tweets_df$lat) ==FALSE)/dim(tweets_df)[1]*100
```

A common NLP data exploration practice is looking at unigram frequencies.
*N-grams* are sequences of N items/tokens. Common n-grams that are used in NLP are unigrams, bigrams, and trigrams. We'll dive deeper into applications of n-grams later (note to self), but we introduce the topic now because it is common to find n-gram frequencies, as we do below with unigrams.
```{r}
head(text_df %>% dplyr::count(word, sort = TRUE),10)
```
We can graph the unigram frequencies with ggplot for a nice visualization of our text data.
```{r}
text_df %>%
  count(word, sort = TRUE) %>%
  filter(n>1700) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) +
  ylab(NULL) +
  coord_flip()
```
Similarly, we can look at common *bigrams*, which are two words in a row.
```{r}
bigrams <- text_df %>%
  unnest_tokens(bigram, word, token = "ngrams", n = 2)

head(bigrams %>%
  count(bigram, sort = TRUE),10)
```


For another fun exploration graphic, try word clouds!
```{r}
text_df %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

##Sentiment
We want to classify the tweets by sentiment, or rather by the feelings and emotions portrayed by the user. As humans we can easily tell if someone is angry, sad, satisfied, or even sarcastic. However we have to give the computer some clues on how to do this start with the `sentiments` dataset which is loaded with the `tidytext` library. Similar to the `stop_words` dataset, there are three lexicons:

* `afinn` gives each word an integer score between -5 and 5, from most negative to most positive sentiment.

* `bing` gives a binary "negative" or "positive" sentiment to each words.

* `nrc` assigns sentiment labels of "positive, negative, anger, anticipation, disgust, fear, joy, sadness, surprise, and trust."

We can also use the tidytext's `get_sentiments()` function to only get data from a particular lexicon.

```{r}
head(sentiments)
head(get_sentiments('afinn'))
```
"How were these sentiment lexicons put together and validated? They were constructed via either crowdsourcing (using, for example, Amazon Mechanical Turk) or by the labor of one of the authors, and were validated using some combination of crowdsourcing again, restaurant or movie reviews, or Twitter data. Given this information, we may hesitate to apply these sentiment lexicons to styles of text dramatically different from what they were validated on, such as narrative fiction from 200 years ago. While it is true that using these sentiment lexicons with, for example, Jane Austen’s novels may give us less accurate results than with tweets sent by a contemporary writer, we still can measure the sentiment content for words that are shared across the lexicon and the text."
(note to self)

Next let's merge some of our data with the sentiment lexicon "afinn" to explore our text.
```{r}
nrc_example <- get_sentiments("afinn") %>% 
  filter(score == -2)

head(text_df %>%
  inner_join(nrc_example) %>%
  count(word, sort = TRUE),10)
```
We note that sentiment analysis would not be straight-forward. From the text printed above, we can't exactly pinpoint if the user is being negative about vaccinations, or the effects of not getting vaccinated, or just being negative about the anti-vaccination or vaccination communities.

Let's plot the average `afinn` scores for each tweet to explore the overall sentiment of our data.

```{r,message=FALSE}
library(tidyr)

tweet_sentiment <- text_df %>%
  anti_join(stop_words) %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(int) %>%
  mutate(text = paste0(word)) %>%
  summarize(sentiment_mean = mean(score))

tweet_sentiment_ordered <- tweet_sentiment[order(tweet_sentiment$sentiment_mean),] 
head(tweet_sentiment_ordered)

tweet_sentiment_ordered$int <- factor(tweet_sentiment_ordered$int, levels = tweet_sentiment_ordered$int)

ggplot(tweet_sentiment_ordered, aes(sentiment_mean)) +
  geom_histogram(show.legend = FALSE, binwidth = .3) +
  ggtitle("Mean afinn sentiment score per tweet") + 
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  ylab(NULL)
 # #facet_wrap(~book, ncol = 2, scales = "free_x")


```

Let's print out some positive tweets. We can use the index column "int" that we've kept constant to print out the original (sans and Twitter handles) tweets.

```{r}
positive <- tweet_sentiment$int[tweet_sentiment$sentiment_mean > 2]
head(tweets_df$text[positive])
```
And let's look at some negative ones as well..
```{r}
negative <- tweet_sentiment$int[tweet_sentiment$sentiment_mean < -2 ]
head(tweets_df$text[negative])
```
Neither the positive nore negative tweets are unanamously on either side of this "issue"", even with this small sample of text. Now we can really see how this won't be easy!

We should think about what we can do about the words that won't be found in the `sentiments` dataset, like "#vaccineswork" that might give us even more information about whether or not the tweet is supporting vaccines or not.

Sentiment may be just a small part of clustering tweets as "pro-vaccination" or "anti-vaccination."

http://www.public.asu.edu/~huanliu/papers/sbp14.pdf

# Data analysis
Throughout our data exploration, we realized the problem we are solving is not quite what we started with. This is not a sentiment analysis: this is classifying tweets as one of two sides of a debate. 

Assuming that, with high probability, users do not change their views--expecially within a 7-day window, we can label Twitter users and represent them as a *social network* to help classify tweets.

To do this, we will use the power of retweets and seed users, or prominent figures that clearly have one opion or another.For example, the World Health Organization and CDC are two entities that would produce pro-vaccination tweets. The data we have also contains reply tweets, but we cannot count on these to identify similar opinion because users may reply to tweets they agree with or tweets they do not agree with. We use Retweets because we can assume that if a user retweets another users tweets about vaccination, they probably agree with their opinion. This assumption gets stronger as the social network grows more connections.

## Label propagation
We will build a network using the `igraph` package. This network will have users as nodes or vertices, with an edge between two vertices if the users retweet the same tweet, are following or are friends with the other user, or retweeted a tweet directly from the other user. The weight of the edge will increase if more than one of the previous are true.


First let's find the proportion of retweets that are exactly the same as the text they retweeted. This is because we don't want to create an edge between two users if one of them added extra text to the retweet, possibly talking about how they disagree with the tweet.

`t1` and `t2` are vectors with either the `retweet_text` or `text` of all the tweets that are retweets, as in the `retweet_text` column is not `NA`.

```{r}
tweets_df <- read.csv("data/date01_14_n18k_key2_uuidgens_final.csv",stringsAsFactors = FALSE)

is_retweet <- is.na(tweets_df$retweet_text) == FALSE

t1 <- tweets_df$retweet_text[is_retweet]
t2 <- tweets_df$text[is_retweet]

sum((t1==t2) == FALSE)/length(t1)*100
```
Only a small percentage of retweets added text. We can safely leave out this data and still have a large dataset.
Let's add a column that labels the retweets as identical or not (or na, if the tweet is not a retweet).
This leaves NAs where `tweets_df$text` is NA.


```{r}
tweets_df$rt_is_identical <- 
  (tweets_df$retweet_text == tweets_df$text)

head(tweets_df$rt_is_identical)
```
We need to build our network. To do this we will use the library `igraph`. First we should create a data frame where the first two columns are vertices with an edge between them. add attributes if there are multiple connections necessary.

```{r}
tweets_df$tweet_factor <- 
  as.numeric(factor(tweets_df$text))

l <- list()

for (i in c(1:max(tweets_df$tweet_factor))){
  users <- unique(c(tweets_df$uuidgens_user[tweets_df$tweet_factor==i],
                    tweets_df$uuidgens_retweet[tweets_df$tweet_factor==i]))
  users <- users[is.na(users) == FALSE]
  l <- append(l, list(users))
}

#namelst <-  c(paste0("tweet",c(1:max(tweets_df$tweet_factor)) , sep = ""))
#names(l) <- namelst
```
Now we have each tweet on the list with the associated users. Now let's create an edge between all these users by creating the dataframe that will be converted to a graph object with `igraph`.

```{r eval=FALSE}
init_test <- data.frame(t(combn(l[[1]],2)))

for (i in c(2:length(l))){
  if (length(l[[i]]) > 1){
    combos <- t(combn(l[[i]],2))
    init <- rbind(init, data.frame(combos))
  }
}
```



So now we've connect each user to all of its tweets, and each unique tweet to all unique users that retweeted it.

```{r}
graph_df <- init %>%
  group_by(X1,X2) %>%
  tally()
```


Add user to user edges
```{r}

```

IMPORT FRIENDS
The column "pro_vax" is a 1 for users who the world health organization follows, and a 0 for users HealthRanger, and anti-vax twitter account, follows. The package rtweet refers to users that a user follows as "friends".
```{r, warning=FALSE}
seeds_df <- read.csv("data/seeds_WHO_HR_friends.csv",stringsAsFactors = FALSE)
head(seeds_df)
seed_users <- seeds_df[(dim(seeds_df)[1] - 1) : dim(seeds_df)[1],]
friends_df <- seeds_df[1:(dim(seeds_df)[1] - 2),]
```
Create `vtx_df` so we can keep track of the order of vertices. Then we can add vertex attributes and the seed users.
```{r, message=FALSE, warning= FALSE}

vtx_df <- data.frame("unique_vtcs" = unique(c(network_df$text, network_df$uuidgens_user)))
vtx_df <- merge(vtx_df, seeds_df, by.x = "unique_vtcs", by.y = "uuidgens_user", all = TRUE)

vtx_df$pro_vax[vtx_df$pro_vax == 0] <- 2

g <- graph_from_data_frame(d = network_df, directed = FALSE, vertices = vtx_df)


randn <- c(3:(sum(is.na(vtx_df$pro_vax) == TRUE)+2))
vtx_df$pro_vax[is.na(vtx_df$pro_vax) == TRUE] <- randn
lpc <- cluster_label_prop(g, initial = vtx_df$pro_vax)
```
Get a random sample and label anti-vax and pro-vax seed users.
```{r, eval=FALSE}
network_df_sample <- network_df[sample(nrow(network_df), 2000), ]

vtx_df <- data.frame("unique_vtcs" = unique(c(network_df_sample$text, network_df_sample$uuidgens_user)))
vtx_df <- merge(vtx_df, seeds_df, by.x = "unique_vtcs", by.y = "uuidgens_user", all = TRUE)

vtx_df$pro_vax[vtx_df$pro_vax == 0] <- 2

g <- graph_from_data_frame(d = network_df_sample, directed = FALSE, vertices = vtx_df)

# Add random labels for remaining vertices (initialized vertices)
randn <- c(3:(sum(is.na(vtx_df$pro_vax) == TRUE)+2))
vtx_df$pro_vax[is.na(vtx_df$pro_vax) == TRUE] <- randn

```

Run label propagation, Add vertex attriutes and color the largest communities.
```{r}
clp <- cluster_label_prop(g, initial = vtx_df$pro_vax)


# create some vertex attributes
V(g)$community <- clp$membership
V(g)$betweenness <- betweenness(g, v = V(g), directed = F)
#V(g)$degree <- degree(g, v = V(g))


data <- toVisNetworkData(g)
nodes <- data[[1]]
edges <- data[[2]]

library(RColorBrewer)

get_big_coms <- nodes %>% 
  select(community) %>% 
  group_by(community) %>% 
  tally() %>%
  arrange(desc(n))

biggest_groups<- get_big_coms$community[1:2]

col <- rep("grey", length(V(g)))
colors<- brewer.pal(10, "Paired")

for(j in c(1:2)){
  col[V(g)$community == biggest_groups[j]] <- colors[j]
}
#graphjs(g, vertex.color = col,  vertex.size = .3, na.ok = TRUE)

```

Let's play around with the vertex attributes in a data frame. We can look at the pro-vax and anti-vax labels and see what communities they ended up in. Then let's color those communities.
```{r}
vtxattr_df <- data.frame(vertex_attr(g))
#head(vtxattr_df)
vtxattr_df %>% filter(pro_vax == 1) %>% select(community) %>% group_by(community) %>% tally()
vtxattr_df %>% filter(pro_vax == 2) %>% select(community) %>% group_by(community) %>% tally()


pro_communities<- vtxattr_df %>% filter(pro_vax == 1) %>% select(community) %>% group_by(community)# %>% tally()
anti_communities<- vtxattr_df %>% filter(pro_vax == 2) %>% select(community) %>% group_by(community)# %>% tally()

vtxattr_df %>% filter(community == 47) %>% tally()
vtxattr_df %>% filter(community == 46) %>% tally()
```
```{r}
library(RColorBrewer)

col <- rep("grey", length(V(g)))
colors<- brewer.pal(10, "Paired")

for(i in pro_communities){
  col[V(g)$community == i] <- colors[1]
}

for(i in anti_communities){
  col[V(g)$community == i] <- colors[2]
}

#graphjs(g, vertex.color = col,  vertex.size = .3, na.ok = TRUE)
```


# Summary of results
```{r}
labeled <- 100*(1 - sum(col == "grey")/length(col))
pro_labeled <- 100*sum(col == colors[1])/length(col)
anti_labeled <- 100*sum(col == colors[2])/length(col)
```
About `r labeled` 10% of our data is labeled!

`r pro_labeled` is pro vax while `r anti_labeled` is antivax

```{r}
vtxattr_df$colors <- col
new_provax <- vtxattr_df %>% filter(is.na(pro_vax)) %>% filter(colors == col[2]) %>% select(name)
new_provax[sample(nrow(new_provax), 10), ]
```



